Which Scenario Should We Design For? Insights from House Elevation for the Multiple PDF Problem

### 2022-01-27 papers/paper/preprint.tex
Track some suggestions from Klaus

% footer
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage[en-US]{datetime2}
\fancyhf{}
\fancyhead[L]{CONFIDENTIAL DRAFT -- Doss-Gollin \& Keller}
\fancyhead[R]{\DTMnow}
\fancyfoot[R]{page~\thepage~of~\pageref{LastPage}}
\pagestyle{fancy}

@@

% authors
\usepackage{authblk}
\title{Which Scenario Should We Design For? Insights from House Elevation for the Multiple PDF Problem}
\author[1]{James Doss-Gollin}
\author[2,3,4]{Klaus Keller}
\affil[1]{Department of Civil and Environmental Engineering, Rice University}
\affil[2]{Department of Geosciences, the Pennsylvania State University}
\affil[3]{Earth and Environmental Systems Institute, the Pennsylvania State University}
\affil[4]{Thayer School of Engineering, Dartmouth College}
\renewcommand*{\Affilfont}{\normalsize\normalfont}

% ACRONYMS
\usepackage[acronym,nopostdot,nonumberlist,shortcuts,]{glossaries}
\newacronym{bdt}{BDT}{Bayesian decision theory}
\newacronym{bfe}{BFE}{base flood elevation}
\newacronym{fema}{FEMA}{the Federal Emergency Management Agency}
\newacronym{gcm}{GCM}{general circulation model}
\newacronym[]{gev}{GEV}{generalized extreme value}
\newacronym{lsl}{LSL}{local mean sea level}
\newacronym{mcmc}{MCMC}{Markov Chain Monte Carlo}
\newacronym{pdf}{PDF}{probability density function}
\newacronym{rcp}{RCP}{representative concentration pathway}
\newacronym{slr}{SLR}{relative sea level rise}
\newacronym{rdm}{RDM}{robust decision making}
\newacronym[plural=SOWs,descriptionplural=states of the world]{sow}{SOW}{state of the world}
\newacronym{tc}{TC}{tropical cyclone}

@@

% use biblatex
\usepackage{csquotes}
\usepackage[
  backend=biber,
  doi=true,
  url=false,
  isbn=false,
  style=authoryear-comp,
  natbib=true,
  backref=false,
  maxbibnames=10,
  maxcitenames=2,
  uniquename=false,
  uniquelist=false,
  sorting=nyt,
]{biblatex}
\renewbibmacro{in:}{}
\AtEveryBibitem{\clearfield{month}\clearfield{day}\clearfield{pages}\clearlist{language}}
\addbibresource{library.bib}

@@

\begin{abstract}
    TBD
\end{abstract}

@@

Since infrastructure built today is likely to persist for several decades, infrastructure designs must take into account the stresses that a project is likely to face far into the future.
On the one hand, overly optimistic assumptions of future risk can lead to unacceptably high risk of failure in the future if additional investments are not made.
For example, a failure to prepare Texas's interconnected natural gas, electricity, and water systems for temperatures that had been observed in the past 30 years caused cascading failures with disparate impacts \citep{doss-gollin_txtreme:2021,busby_cascadingrisks:2021}.
On the other, unless opportunity costs are negligible, designing for a too-pessimistic scenario can contribute to maladptation by limiting resource availability in the future.
For example, \ldots\james{add an example of a maladaptive megaproject -- look to sustainability science or the arguments in \citet{ansar_bigisfragile:2017}, something by Flyvbjerg, \etc}

TODO: set up flood risk management as a context where over- or under-design could be relevant, then raise house elevation in the coastal zone as a specific example.
Federal guidance for this decision relies on standards (the $T$-year flood) rather than risk-based design.
Be sure to note that this has parallels to lots of other engineering design guidance. Then note the work of \citet{zarekarizi_suboptimal:2020} or \citet{xian_elevation:2017}, who show that a risk-based approach may improve significantly on a standards-based approach.

@@

This suggests that either nonstationarity should be ignored, or subjective approaches for designing for nonstationarity should be adapted.
Given that physical theories consistent with the observational record and models robustly project changes to climate hazard under warning, ignoring nonstationarity seems suboptimal.
Thus decision-makers need some subjective way to choose which possible futures to design for.

In this paper, we use a didactic case study of house elevation in Norfolk, VA as a lens through which to shed light on the following research questions.
\begin{description}
    \item[Q1] What are the potential benefits to homeowners of changing threshold-based guidance to risk-based guidance under nonstationary flood hazard?
    \item[Q2] How does risk-based guidance change under different assumptions of physical model structure and human emissions forcings?
    \item[Q3] How can deep uncertainties be transparently and consistently synthesized for decision analysis?
\end{description}
Specifically, we examine this problem through three lenses.

The paper is organized as follows.
In \cref{sec:case-study} we describe our case study in Norfolk, VA.
First, in \cref{sec:multiple-simulation} we analyze each simulation (\ie, realizations from these \glspl{pdf}) independently and illustrate .
Next, in \cref{sec:multiple-pdf} we consider the design problem as a set of ``multiple \glspl{pdf}'' and use these findings to motivate theoretical advances.
Then in \cref{sec:synthesizing} we present a probabilistic method for synthesizing simulations from multiple \glspl{pdf} and illustrate the advantages and potential pitfalls of this approach.
Finally in \cref{sec:conclusions} we briefly summarize our findings, discuss implications for engineering practice, and identify needs for future research.

@@

We model a \emph{hypothetical} house in Norfolk, VA, where sea level rise drives nonstationary future flood hazard.

For interpretability, we focus on deep uncertainty in \gls{slr} and treat other model parameters as fixed quantities or as outputs from a single probabilsitic model, as shown in \cref{tab:uncertainties}.
We consider a single, one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}).
We evaluate this decision for each of many possible simulations of \gls{slr} using a combined engineering-economic model based on that of \citet{zarekarizi_suboptimal:2020}.
In the remainder of this section we discuss data sources and the system model.

\begin{table}
    \centering
    \begin{tabular}{l l p{3in} l}
        \toprule
        Name           & Symbol           & Conceptual framework                                                                 & Section               \\
        \midrule
        \Gls{slr}      & $\overline{y}_t$ & Deeply uncertain: multiple methods considered                                        & \ref{sec:sea-level}   \\
        Storm surge    & $y'_t$           & Probabilistic: a single stationary \acrshort{gev} model                              & \ref{sec:storm-surge} \\
        Discount rate  & $1-\gamma$       & Fixed at 2\%                                                                         & \ref{sec:ead}         \\
        Depth-damage   & $D(h-y_t)$       & Deterministic: based on HAZUS model                                                  & \ref{sec:ead}         \\
        Elevation cost & $C(\Delta h)$    & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} & \ref{sec:sow-metrics} \\
        Initial height & $h_0$            & Deterministic: \SI{0.5}{ft} above \gls{bfe} unless otherwise noted                   & TBD                   \\
        \bottomrule
    \end{tabular}
    \caption{Summary of uncertainties modeled and their treatment.}\label{tab:uncertainties}
\end{table}

@@

\Cref{fig:msl-boxplots} shows the \glspl{pdf} of mean sea level in 2100 under each of the eight models considered: four emissions pathways (\gls{rcp} scenarios) times two models of ice sheet dynamics.
Variation between different \glspl{pdf} is dominated by the \gls{rcp} 8.5 scenario with fast ice sheet dynamics \citep[see][for a discussion of these dynamics]{wong_brick0.2:2017}.
This is also the \gls{pdf} with the greatest within-model variance.
\begin{figure}
    \centering
    \includegraphics[width=5in]{msl_boxplots.pdf}
    \caption{
        Sea level rise projections use the BRICK model \citep{wong_brick0.2:2017,ruckert_coastal:2019}.
        We consider \emph{eight time-dependent PDFs of sea level rise}: four \gls{rcp} scenarios $\times$ two parameterizations of ice sheet processes.
    }\label{fig:msl-boxplots}
\end{figure}

@@

Data on storm surge comes from Sewells Point, VA (NOAA gauge 8638610).\james{cite}
Hourly recordings of water level are available from 1928 to the present; we use data from the period January 01 1928 to December 31 2021.
For each year we remove the annual mean then extract a single maximum; we refer the set of these maxima for all years as the time series of annual maximum storm surges.
This time series is shown in \cref{fig:observed-surges}(a).\james{This figure needs a bit of work: subpanel plots, equalizing axes, narrower time series, and no lines between the points}

\begin{figure}
    \centering
    \includegraphics[height=1.5in]{historic_surge.pdf}%
    \includegraphics[height=1.5in]{surge_return_period.png}
    \caption{
        Annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA.
        (a):
        time series of historic storms.
        Purple (orange) arrows denote notable tropical cyclones (nor'easters).
        (b):
        return periods.
        Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull formula .
        Blue line shows a maximum likelihood GEV fit and the gray lines show Monte Carlo draws from the full posterior distribution of the Bayesian GEV fit.
    }
    \label{fig:observed-surges}
\end{figure}

We model future storm surge risk through a probabilistic lens.
To do this, we develop a stationary Bayesian model for storm surge:
\begin{equation}\label{eq:surge-model}
    y'_t \sim \text{GEV}\left(\mu, \sigma, \xi \right),
\end{equation}
where $y'_t$ is the storm surge (above sea level) in year $t$.

When GEV distributions are fit to time series data without strong prior information, the resulting uncertainties can be unrealistically large.\james{cite}
Thus, information from other locations, time periods, and/or from physical principles may help to reduce uncertainties \citep{Merz:2008bl,Merz:2008eh}.
To constrain estimates, we add weakly informative priors.
However, rather than adding prior information over the parameters themselves, we follow \citet{coles_evd:1996} and apply a prior over extreme quantiles of the distribution.
Specifically, we apply LogNormal priors over the 90th, 99th, and 99.9th percentiles of the distribution (\ie, the 10, 100, and 1000 year return period events).
These values were chosen to represent plausible physical ranges (the assumed 90\% confidence interval for the 10 year event is XY years, for the 100 year event is XZ years, and for the 1000 year event ZY years) and are plotted in supporting figure SX.\james{return to this sentence}

@@

The expected annual damages is sometimes calculated by assuming analytically tractable functional forms for the depth-damage relationship and for the  distribution of hazard \citep[\eg][]{vandantzig_dike:1956}.
However, the convolution of the HAZUS depth-damage equation with the GEV posterior does not have a tractable analytic solution so we instead estimate it through a Monte Carlo method:
\begin{enumerate}
    \item For $k=1, \ldots, K$:
          \begin{enumerate}
              \item sample a draw from the posterior distribution of flood hazard to get $\left\{ \mu_k, \sigma_k, \xi_k \right\}$
              \item simulate a single storm surge from the stationary distribution (equation XYZ) and add the mean sea level to get total flood depth
              \item calculate the flood damages for this draw by plugging the annual maximum flood depth ($h - y_t$) into  the deterministic HAZUS depth-damage relationship
              \item store this as the $k$th damage
          \end{enumerate}
    \item Estimate expected annual damages as the sample mean of the $K$ estimates
\end{enumerate}

@@

The second component of the system model converts a time series of annual expected damages (\cref{sec:ead}) into lifetime expected damages, which we define as the upfront cost plus the discounted sum of expected annual damages
\begin{equation}\label{eq:led}
    \textrm{LED} = \sum_{t=1}^T \gamma^{t-1} \mathbb{E}[d_t | \Delta h],
\end{equation}
where $\gamma$ is the 1 minus the discount rate.
For the didactic purposes of this study we neglect uncertainty in the discount rate \citep{zarekarizi_suboptimal:2020,arrow_discount:2013} and use a fixed discount rate of 2, thus $\gamma = 0.98$.
Neglecting uncertainty in house lifetime, we model flood damages from 2022 to 2071, thus $T=50$.\james{This has not been the case thus far}
We calculate lifetime expected damages \emph{for each \gls{sow} separately}.

@@

\begin{figure}
    \centering
    \includegraphics[width=4in]{bayes-rdm.pdf}
    \caption{
        Outline of the generally applicable decision theoretic framework developed.
        Boxes are labeled in the order in which they are presented in the text: we begin in \cref{sec:multiple-simulation} using a lens of exploratory modeling, \ie boxes  (a), (b), and (c), to analyze the elevation problem.
        In \cref{sec:multiple-pdf} we discuss Boxes (d) and (e) to understand scenario-conditional probability distributions and the ``multiple \acrshort{pdf} problem''.
        Finally in \cref{sec:synthesizing} we add boxes (f) and (g) to synthesize across multiple \glspl{pdf}.
    }\label{fig:flowchart}
\end{figure}

@@

The literature on decision making under deep uncertainty \citep[see][]{marchau:2019} offers some insight as to how to identify decisions that perform well over many plausible \acrfullpl{sow}.
The first that we consider is exploratory modeling, which sheds light the questions: ``how well does a given decision perform under different possible futures, and to which possible futures are different decisions vulnerable?''

\Cref{fig:scenario-map} shows each possible future as a single point for each of several values of $\Delta h$.\klaus{Would it be better here to have a figure that has $\Delta h$ on the $y$ axis and uses expected lifetime cost to color the points?}
On the $y$ axis is expected lifetime costs (see \cref{sec:sow-metrics}).
On the $x$ axis is mean sea level in 2100, which is an interpretable 1-dimensional projection of the \glspl{sow}.
Although our model for expected discounted total costs is deterministic, points with similar $x$-values may have different $y$-values because their sea levels before 2100 can differ (\ie, the 1-dimensional projection of the \glspl{sow} does not represent all variance between them).

\begin{figure}
    \includegraphics[width=\textwidth]{scenario_maps.png}
    \caption{
        Scenario maps: one dot represents one \gls{sow}.
        Some work is still needed -- I think that changing to four values of $\Delta h$ and showing two or three different values of $h_0$ might be useful (alternatively we could vary the discount rate $\gamma$).
        Also need to rescale the $y$-axis by house value.
        It would also be cool to create an interactive online version of this figure that lets you scroll over a particular dot and learn which scenario it's from!
        %TODO
    }\label{fig:scenario-map}
\end{figure}

@@

An alternative interpretation of the available \glspl{sow} is as IID draws from each of eight \glspl{pdf} of sea level rise, as shown in \cref{fig:msl-boxplots}.
This has the approach of a probabilistic interpretation: \emph{conditional on a particular model} (\gls{rcp} scenario and ice sheet model structure), we can view corresponding \glspl{sow} as IID draws from the distribution of outcomes.

A more formal notation is introduced in boxes (d) and (e) of \cref{fig:flowchart}.
Specifically, for each of $K$ models (box d), which here are our eight models for \gls{slr},\james{need to go through whole document and be consistent about MSL, SLS, SRL, RSLR, \etc} the samples of metrics are taken to be IID samples from the distribution of outcomes (box e).
Since we define these models to estimate $p(\Omega)$, our models $M_1, \ldots, M_K$ return a probability of $1 / J_K$ for each of the $J_K$ simulations from the $K$th model, and $0$ for simulations from all other models.
We can then use this conditional distribution over outcomes to create metrics that depend only on the decision, such as averages or quantiles of outcomes.

\begin{figure}
    \centering
    \includegraphics[width=2.5in]{tradeoffs_height_totalcost_byscenario.pdf}~
    \includegraphics[width=2.5in]{tradeoffs_height_iqr_byscenario.pdf}
    \caption{
        Tradeoffs between (L) construction cost and expected lifetime costs or (R) construction cost and the uncertainty of future lifetime costs depend on the PDF selected.
    }
    \label{fig:tradeoffs}
\end{figure}

@@

Motivated by the multiple \gls{pdf} problem, we develop in this section a framework for synthesizing information across \glspl{pdf} using subjective Bayesian modeling.

Specifically, we frame the set of \glspl{sow} through a sampling lens and consider the possibility that the distribution from which they are sampled, $p_\textrm{empirical}(\Omega)$, does not reflect our belief about their true distribution, $p_\mathrm{true}(\Omega)$.
Often this is by design: exploratory modeling emphasizes the value of sampling low-probability futures as a way to learn about the system and develop insight \citep{bankes:1993}.
This is analagous to the deliberate over- or under-sampling of certain groups in survey design, and so following standard practice in survey design we develop a re-weighting approach to correct for the over- or under-sampling of different regions of the survey space.

To calculate these weights, we apply a grid-based discretization scheme.
First, we project the \glspl{sow} $s_1, \ldots, s_J$ onto a one-dimensional space, which we denote $\psi_1, \ldots, \psi_J$.
As before, we take this space to be the mean sea level in 2100.
Without loss of generality, we assume the $\psi_j$ to be sorted from least to greatest so that $\psi_{j-1} \leq \psi_j$, ($j \neq 1$).
This allows us to simplify $p_\mathrm{true}(s_j) = p_\mathrm{true}(\psi_j)$.\klaus{This notation seems bad, suggestions?}
Defining $F_\mathrm{true}(s)$ to be the cumulative distribution corresponding to $p_\mathrm{true}$, \ie
$$F_\mathrm{true}(s) = \int_{-\infty}^{s'} p_\mathrm{true}(s) \dd{s},$$
we calculate weights using \cref{eq:weights} as illustrated in \cref{fig:weighting}.
\begin{equation}\label{eq:weights}
    w_j = \begin{cases}
        j = 1     & F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_1 + \psi_2])                                                                   \\
        1 < j < J & F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_{j} + \psi_{j+1}]) - F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_{j-1} + \psi_j]) \\
        j = J     & 1 - F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_{J-1} + \psi_J]).
    \end{cases}
\end{equation}
This grid-based weighting scheme can be used with higher-dimensional projections of the scenarios, as is done for example in the joint probability method used for hurricanes \citep[see, \eg,][]{johnson_clara:2013,toro_jpm-os:2010}.

>>>>

% footer
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage[en-US]{datetime2}
\fancyhf{}
\fancyhead[L]{CONFIDENTIAL DRAFT by J. Doss-Gollin \& K. Keller}
\fancyhead[R]{\DTMnow}
\fancyfoot[R]{page~\thepage~of~\pageref{LastPage}}
\pagestyle{fancy}

@@

% authors
\usepackage{authblk}
\title{Which Scenario Should We Design For? Insights from House Elevation for the Multiple PDF Problem}
\author[1]{James Doss-Gollin}
\author[2]{Klaus Keller}
\affil[1]{Department of Civil and Environmental Engineering, Rice University}
\affil[2]{Thayer School of Engineering, Dartmouth College}
\renewcommand*{\Affilfont}{\normalsize\normalfont}

% ACRONYMS
\usepackage[acronym,nopostdot,nonumberlist,shortcuts,]{glossaries}
\newacronym{bdt}{BDT}{Bayesian decision theory}
\newacronym{bfe}{BFE}{base flood elevation}
\newacronym{fema}{FEMA}{the Federal Emergency Management Agency}
\newacronym{gcm}{GCM}{general circulation model}
\newacronym[]{gev}{GEV}{generalized extreme value}
\newacronym{iid}{IID}{independent and identically distributed}
\newacronym{lsl}{LSL}{local mean sea level}
\newacronym{mcmc}{MCMC}{Markov Chain Monte Carlo}
\newacronym{pdf}{PDF}{probability density function}
\newacronym{rcp}{RCP}{representative concentration pathway}
\newacronym{slr}{SLR}{relative sea level rise}
\newacronym{rdm}{RDM}{robust decision making}
\newacronym[plural=SOWs,descriptionplural=states of the world]{sow}{SOW}{state of the world}
\newacronym{tc}{TC}{tropical cyclone}

@@

% use biblatex
\usepackage{csquotes}
\usepackage[
  backend=biber,
  doi=true,
  url=false,
  isbn=false,
  style=authoryear-comp,
  natbib=true,
  backref=false,
  maxbibnames=10,
  maxcitenames=2,
  uniquename=false,
  uniquelist=false,
  sorting=nyt,
  giveninits=true,
]{biblatex}
\renewbibmacro{in:}{}
\AtEveryBibitem{\clearfield{month}\clearfield{day}\clearfield{pages}\clearlist{language}}
\addbibresource{library.bib}

@@

\begin{abstract}
    \textbf{Draft in preparation to Earth's Future}.
\end{abstract}

@@

Since infrastructure built today is likely to persist for several decades, infrastructure designs must take into account the stresses that a project is likely to face decades into the future.
On the one hand, overly optimistic assumptions of future risk can lead to unacceptably high risk of failure in the future if additional investments are not made.
For example, a failure to prepare Texas's interconnected natural gas, electricity, and water systems for temperatures that had been observed in the past 30 years caused cascading failures with disparate impacts \citep{doss-gollin_txtreme:2021,busby_cascadingrisks:2021}.\james{Too fast on intro. Need to slow down and define concepts}
On the other, unless opportunity costs are negligible, designing for a too-pessimistic scenario can contribute to maladptation by limiting resource availability in the future.
For example, \ldots\james{add an example of a maladaptive megaproject -- look to sustainability science or the arguments in \citet{ansar_bigisfragile:2017}, something by Flyvbjerg, \etc}

TODO: set up flood risk management as a context where over- or under-design could be relevant, then raise house elevation in the coastal zone as a specific example.
Federal guidance for this decision relies on standards (the $T$-year flood) rather than risk-based design.
Be sure to note that this has parallels to lots of other engineering design guidance. Then note the work of \citet{zarekarizi_suboptimal:2020} or \citet{xian_elevation:2017}, who show that a risk-based approach may improve considerably on a standards-based approach.

@@

This motivates subjective approaches for designing for nonstationarity.
Thus decision-makers need some subjective way to choose which possible futures to design for.

In this paper, we use a didactic case study of house elevation in Norfolk, VA as a lens through which to shed light on the following research questions.
\begin{description}
    \item[Q1] What are the potential benefits to homeowners of changing threshold-based guidance to risk-based guidance under nonstationary flood hazard?
    \item[Q2] How does risk-based guidance change under different assumptions of physical model structure and human emissions forcings?
    \item[Q3] How can deep uncertainties be transparently and consistently synthesized for decision analysis?
\end{description}
Specifically, we examine this problem through three lenses.\james{Which ones?}

The paper is organized as follows.\james{This is confusing to the reader}
In \cref{sec:case-study} we describe our case study in Norfolk, VA.
First, in \cref{sec:multiple-simulation} we analyze each simulation (\ie, realizations from these \glspl{pdf}) independently and illustrate .
Next, in \cref{sec:multiple-pdf} we consider the design problem as a set of ``multiple \glspl{pdf}'' and use these findings to motivate theoretical advances.
Then in \cref{sec:synthesizing} we present a probabilistic method for synthesizing simulations from multiple \glspl{pdf} and illustrate the advantages and potential pitfalls of this approach.
Finally in \cref{sec:conclusions} we briefly summarize our findings, discuss implications for engineering practice, and identify  future research needs.

@@

We model a \emph{hypothetical} house in Norfolk, VA, where sea level rise drives nonstationary future flood hazard.
For interpretability, we focus on deep uncertainty in \gls{slr} and treat other model parameters as fixed quantities or as outputs from a single probabilsitic model, as shown in \cref{tab:uncertainties}.
We consider a single, one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}).
We evaluate this decision for each of many possible simulations of \gls{slr} using a combined engineering-economic model based on that of \citet{zarekarizi_suboptimal:2020}.
In the remainder of this section we discuss data sources and the system model.

\begin{table}
    \centering
    \caption{Summary of uncertainties modeled and their treatment.}\label{tab:uncertainties}
    \begin{tabular}{l l p{3in} l}
        \toprule
        Name           & Symbol           & Conceptual framework                                                                 & Section               \\
        \midrule
        \Gls{slr}      & $\overline{y}_t$ & Deeply uncertain: multiple methods considered                                        & \ref{sec:sea-level}   \\
        Storm surge    & $y'_t$           & Probabilistic: a single stationary \acrshort{gev} model                              & \ref{sec:storm-surge} \\
        Discount rate  & $1-\gamma$       & Fixed at 2\%                                                                         & \ref{sec:ead}         \\
        Depth-damage   & $D(h-y_t)$       & Deterministic: based on HAZUS model                                                  & \ref{sec:ead}         \\
        Elevation cost & $C(\Delta h)$    & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} & \ref{sec:sow-metrics} \\
        Initial height & $h_0$            & Deterministic: \SI{0.5}{ft} above \gls{bfe} unless otherwise noted                   & TBD                   \\
        \bottomrule
    \end{tabular}
\end{table}

@@

\Cref{fig:msl-boxplots} shows the \glspl{pdf} of mean sea level in 2100 under each of the eight models considered: four emissions pathways \citep[\gls{rcp} scenarios][]{moss_rcp:2010} and two models of ice sheet dynamics.
Variation between different \glspl{pdf} is dominated by the \gls{rcp} 8.5 scenario with fast ice sheet dynamics \citep[see][for a discussion of these dynamics]{wong_brick0.2:2017}.
This is also the \gls{pdf} with the greatest within-model variance.
\begin{figure}
    \centering
    \includegraphics[width=5in]{msl_boxplots.pdf}
    \caption{
        Sea level rise projections use the BRICK model \citep{wong_brick0.2:2017,ruckert_coastal:2019}.
        We consider \emph{eight time-dependent PDFs of sea level rise}: four \gls{rcp} scenarios $\times$ two parameterizations of ice sheet processes.
    }\label{fig:msl-boxplots}
\end{figure}

@@

Data on storm surge comes from Sewells Point, VA (NOAA gauge 8638610).\james{cite NOAA data source}
Hourly recordings of water level are available from 1928 to the present; we use data from the period January 01 1928 to December 31 2021.
For each year we remove the annual mean then extract a single maximum; we refer the set of these maxima for all years as the time series of annual maximum storm surges.
This time series is shown in \cref{fig:observed-surges}(a).
In (b) we use the Weibull plotting position
\begin{equation}\label{eq:plot-pos}
    \frac{r}{N + 1}
\end{equation}
where $N$ is the sample size and $r$ is the order of the $N$ observations ($r=1$ is the largest, $r=N$ is the smallest).

\begin{figure}
    \centering
    \includegraphics[height=1.5in]{historic_surge.pdf}%
    \includegraphics[height=1.5in]{surge_return_period.png}
    \caption{
        Annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA (CITE SOURCE).
        (a):
        Time series of historic storms.
        Purple (orange) arrows denote notable tropical cyclones (nor'easters).
        (b):
        Return periods.
        Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull formula (eq.~\cref{eq:plot-pos}).
        Blue line shows a maximum likelihood \gls{gev} fit and the gray lines show Monte Carlo draws from the full posterior distribution of the Bayesian GEV fit.
    }
    \label{fig:observed-surges}
\end{figure}

We model future storm surge risk through a probabilistic lens.
To do this, we develop a stationary\james{discuss / validate this -- it's cool, we have the supplemental info and can emphpasize our focus here is on SLR for didactic purposes} Bayesian model for storm surge:
\begin{equation}\label{eq:surge-model}
    y'_t \sim \text{GEV}\left(\mu, \sigma, \xi \right),
\end{equation}
where $y'_t$ is the storm surge (above sea level) in year $t$ and $\mu$, $\sigma$, and $\xi$ are the location, scale, and shape parameters, respectively.

When \gls{gev} distributions are fit to time series data without strong prior information, the resulting uncertainties can be unrealistically large.\james{cite}
Thus, information from other locations, time periods, and/or from physical principles may help to reduce uncertainties \citep{Merz:2008bl,Merz:2008eh}.
To constrain estimates, we add weakly informative priors.
However, rather than adding prior information over the parameters themselves, we follow \citet{coles_evd:1996} and apply a prior over extreme quantiles of the distribution.
Specifically, we apply LogNormal priors over the 90th, 99th, and 99.9th percentiles of the distribution (\ie, the 10, 100, and 1000 year return period events).\james{Expand on this ``probabilistic inversion''}
These values were chosen to represent plausible physical ranges (the assumed 90\% confidence interval for the 10 year event is XY years, for the 100 year event is XZ years, and for the 1000 year event ZY years) and are plotted in supporting figure SX.\james{return to this sentence}

@@

The expected annual damages is sometimes calculated by assuming analytically tractable functional forms for the depth-damage relationship and for the  distribution of hazard \citep[\eg][]{vandantzig_dike:1956}.
However, the convolution of the HAZUS depth-damage equation with the \gls{gev} posterior does not have a tractable analytic solution so we instead estimate it through a Monte Carlo method:
\begin{enumerate}
    \item For $k=1, \ldots, K$:
          \begin{enumerate}
              \item sample a draw from the posterior distribution of flood hazard to get $\left\{ \mu_k, \sigma_k, \xi_k \right\}$
              \item simulate a single storm surge from the stationary distribution (equation XYZ) and add the mean sea level to get total flood depth
              \item calculate the flood damages for this draw by plugging the annual maximum flood depth ($h - y_t$) into  the deterministic HAZUS depth-damage relationship
              \item store this as the $k$th damage
          \end{enumerate}
    \item Estimate expected annual damages as the sample mean of the $K$ estimates
\end{enumerate}

@@

The second component of the system model converts a time series of annual expected damages (\cref{sec:ead}) into lifetime expected damages, which we define as the upfront cost plus the discounted sum of expected annual damages
\begin{equation}\label{eq:led}
    \textrm{LED} = \sum_{t=1}^T \gamma^{t-1} \mathbb{E}[d_t | \Delta h],
\end{equation}
where the discount rate is $1 - \gamma$.
For the didactic purposes of this study we neglect uncertainty in the discount rate \citep{zarekarizi_suboptimal:2020,arrow_discount:2013} and use a fixed discount rate of 2\%, thus $\gamma = 0.98$.
Neglecting uncertainty in house lifetime, we model flood damages from 2022 to 2071, thus $T=50$.\james{This has not been the case thus far}
We calculate lifetime expected damages \emph{for each \gls{sow} separately}.

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{bayes-rdm.pdf}
    \caption{
        Outline of the proposed decision theoretic framework.
        Boxes are labeled in the order in which they are presented in the text.
        We begin in \cref{sec:multiple-simulation} using a lens of exploratory modeling, \ie boxes  (a), (b), and (c), to analyze the elevation problem.
        In \cref{sec:multiple-pdf} we discuss Boxes (d) and (e) to understand scenario-conditional probability distributions and the ``multiple \acrshort{pdf} problem''.
        Finally in \cref{sec:synthesizing} we add boxes (f) and (g) to synthesize across multiple \glspl{pdf}.
    }\label{fig:flowchart}
\end{figure}

@@

The literature on decision making under deep uncertainty \citep[see][]{marchau:2019} offers some insight as to how to identify decisions that perform well over many plausible \glspl{sow}.
We first consider exploratory modeling, which sheds light on the questions: ``how well does a given decision perform under different possible futures, and to which possible futures are different decisions vulnerable?''\james{Maybe this doesn't need to be a questino}

\Cref{fig:scenario-map} shows each possible future as a single point for each of several values of $\Delta h$.\james{Add after updating this figure}
On the $y$ axis is expected lifetime costs (see \cref{sec:sow-metrics}).
On the $x$ axis is mean sea level in 2100, which is an interpretable 1-dimensional projection of the \glspl{sow}.
Although our model for expected discounted total costs is deterministic, points with similar $x$-values may have different $y$-values because their sea levels before 2100 can differ (\ie, the 1-dimensional projection of the \glspl{sow} does not represent all variance between them).

\begin{figure}
    \includegraphics[width=\textwidth]{scenario_maps.png}
    \caption{
        Scenario maps showing the dependence of average lifetime cost as a function of \gls{slr} in 2100 for several values of $\Delta h$.
        One dot represents one \gls{sow}.
        %TODO
    }\label{fig:scenario-map}
\end{figure}

@@

An alternative interpretation of the available \glspl{sow} is as \gls{iid} draws from each of eight \glspl{pdf} of sea level rise, as shown in \cref{fig:msl-boxplots}.
This has the approach of a probabilistic interpretation: \emph{conditional on a particular model} (\gls{rcp} scenario and ice sheet model structure), we can view corresponding \glspl{sow} as \gls{iid} draws from the distribution of outcomes.

A more formal notation is introduced in boxes (d) and (e) of \cref{fig:flowchart}.
Specifically, for each of $K$ models (box d), which here are our eight models for \gls{slr},\james{need to go through whole document and be consistent about MSL, SLS, SRL, RSLR, \etc} the samples of metrics are taken to be \gls{iid} samples from the distribution of outcomes (box e).
Since we define these models to estimate $p(\Omega)$, our models $M_1, \ldots, M_K$ return a probability of $1 / J_K$ for each of the $J_K$ simulations from the $K$th model, and $0$ for simulations from all other models.
We can then use this conditional distribution over outcomes to create metrics that depend only on the decision, such as averages or quantiles of outcomes.

\begin{figure}
    \centering
    \includegraphics[width=2.5in]{tradeoffs_height_totalcost_byscenario.pdf}~
    \includegraphics[width=2.5in]{tradeoffs_height_iqr_byscenario.pdf}
    \caption{
        Need a summary sentence.
        %TODO
        Trade-offs between (a) construction cost and expected lifetime costs or (b) construction cost and the uncertainty of future lifetime costs depend on the PDF selected.
    }
    \label{fig:tradeoffs}
\end{figure}

@@

Motivated by the multiple \gls{pdf} problem, we develop in this section a framework for synthesizing information across \glspl{pdf} using subjective Bayesian modeling.\james{This also needs to go in the abstract}

Specifically, we frame the set of \glspl{sow} through a sampling lens and consider the possibility that the distribution from which they are sampled, $p_\textrm{empirical}(\Omega)$, does not reflect our belief about their true distribution, $p_\mathrm{true}(\Omega)$.
Often this is by design: exploratory modeling emphasizes the value of sampling low-probability futures as a way to learn about the system and develop insight \citep{bankes:1993}.\james{Tighten up this writing}
This is analagous to the deliberate over- or under-sampling of certain groups in survey design, and so following standard practice in survey design we develop a re-weighting approach to correct for the over- or under-sampling of different regions of the survey space.

We apply a grid-based discretization scheme to calculate these weights.
First, we project the \glspl{sow} $s_1, \ldots, s_J$ onto a one-dimensional space, which we denote $\psi_1, \ldots, \psi_J$.
As before, we take this space to be the mean sea level in 2100.
Without loss of generality, we assume the $\psi_j$ to be sorted from least to greatest so that $\psi_{j-1} \leq \psi_j$, ($j \neq 1$).
This allows us to simplify $p_\mathrm{true}(s_j) = p_\mathrm{true}(\psi_j)$.\klaus{This notation seems bad, suggestions?}
Defining $F_\mathrm{true}(s)$ to be the cumulative distribution corresponding to $p_\mathrm{true}$, \ie
$$F_\mathrm{true}(s) = \int_{-\infty}^{s'} p_\mathrm{true}(s) \dd{s},$$
we calculate weights using \cref{eq:weights} as illustrated in \cref{fig:weighting}.
\begin{equation}\label{eq:weights}
    w_j = \begin{cases}
        j = 1     & F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_1 + \psi_2])                                                                   \\
        1 < j < J & F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_{j} + \psi_{j+1}]) - F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_{j-1} + \psi_j]) \\
        j = J     & 1 - F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_{J-1} + \psi_J]).
    \end{cases}
\end{equation}
This grid-based weighting scheme can be used with higher-dimensional projections of the scenarios, as is done for example in the joint probability method used for hurricanes \citep[see, \eg,][]{johnson_clara:2013,toro_jpm-os:2010}.
